### 2022-01-09 README.Rmd
first human commit
- Lots of changes

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ASQ3

<!-- badges: start -->
[![R-CMD-check](https://github.com/nim-ach/ASQ3/workflows/R-CMD-check/badge.svg)](https://github.com/nim-ach/ASQ3/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The purpose of the {ASQ3} package is to provide access to the data collected by the Centro Asistencial Docente e Investigación de la Universidad de Magallanes (CADI-UMAG).

The data were treated, eliminating any sensitive information that could compromise the identity of the participants or linked individuals.

## Installation

You can install the development version of ASQ3 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nim-ach/ASQ3")
```

## Example

This is a basic example which shows you how to use the dataset from your <i class="fab fa-r-project"></i> console.

```{r example}
library(ASQ3) # Load the package

head(dataset[, 1:5], n = 5) # Print the first 5 rows and columns
```

For more information about the dataset, you can consult the package documentation.