### 2022-01-17 README.Rmd
Add umag hyperlink and about section

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ASQ3

<!-- badges: start -->
[![R-CMD-check](https://github.com/nim-ach/ASQ3/workflows/R-CMD-check/badge.svg)](https://github.com/nim-ach/ASQ3/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The purpose of the {ASQ3} package is to provide access to the data collected by the Centro Asistencial Docente e Investigación de la [Universidad de Magallanes](http://www.umag.cl/) (CADI-UMAG).

The data were treated, eliminating any sensitive information that could compromise the identity of the participants or linked individuals. Data preparation process can be seen [here](reference/data-prep.html).

## Installation

You can install the development version of ASQ3 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nim-ach/ASQ3")
```

## Example

This is a basic example which shows you how to use the dataset from your <i class="fab fa-r-project"></i> console.

```{r example}
library(ASQ3) # Load the package

head(dataset[, 1:4], n = 5) # Print the first 5 rows of the first 4 columns
```

For more information about the dataset, you can consult the dataset [documentation](https://nim-ach.github.io/ASQ3/reference/dataset.html).

## About <i class="fab fa-r-project"></i>

R is a high-level statistical programming language used for data analysis, as well as for the development of software tools such as websites, reports, applications, digital books, etc. For more information you can visit the [R-project](https://www.r-project.org/about.html) website.